---
title: "Bees!"
author: "Isha, Jade, Valerie"
date: "Winter 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#library("jpeg")
```

## Packages
We use the tidyverse and skimr for our project.
```{r echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
library(tidyverse)
library(skimr)
library("RColorBrewer")
```

## Load the data
We found our data at the QQQ. Citations.
```{r echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
bee <- 
  readr::read_csv("bee2.csv")
```

##Rename the Data
```{r}
bee <- 
  bee %>% 
  dplyr::rename(
          year        = "Year",
          season      = "Season",
          state       = "State",
          totAnnualLoss = "Total Annual Loss",
          beekeepers = "Beekeepers",
          beekeepersExclusive = "Beekeepers Exclusive to State",
          colonies = "Colonies",
          coloniesExclusive =  "Colonies Exclusive to State" 
          )

names(bee)
```


## How many bee colonies are there in each state?
```{r}
bee %>%
  filter(year=="2016/17") %>%
  ggplot( aes(x=state,y=colonies)) +
  geom_bar(stat = "identity") +
  labs(title = "Colonies Vs state",
       x = "states",
       y = "colonies")+
  theme(plot.title = element_text(face="bold", hjust = 0.5),
        axis.text.x = element_text(angle = 60, hjust = 1),
        axis.text.y = element_text(angle = 90, hjust = 1))+
  scale_fill_brewer(palette = "Dark2")
``` 

##BUT WHY ARE BEE POPULATIONS DESCREASING?
##Let's investigate one posible reason: Beekeepers

##Here's How Many Beekeepers Were Employed by Each State in 2016-2017
```{r}
bee %>% 
  filter(year=="2016/17") %>% 
  ggplot(aes(x=state, y= beekeepers, fill = state))+
  geom_col()+
  labs(title = "Number of Beekeepers per State in 2016-2017",
       x = "State",
       y = "Number of Beekeepers Hired by State")+
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=10),
          legend.position="none",
          plot.title = element_text(lineheight=.8, face="bold"))
```

##Pennsylvania and Virginia Are states with higher bee loss 
- Pennsylviania is ranked #16 in highest total annual bee loss
- Virginia is ranked #19
```{r}
bee %>% 
  filter(year=="2016/17") %>% 
  arrange(desc(totAnnualLoss))
```

##There is no relationship between number of colonies and beekeepers
```{r}
bee %>% 
  filter(year=="2016/17") %>% 
  ggplot(aes(x=beekeepers, y=totAnnualLoss, color = state))+
  geom_jitter()+
  geom_smooth(method=lm, se=FALSE)+
  labs(title = "Total Annual Loss vs Number of Beekeepers in 2016-2017",
       x = "Number of Beekeepers Hired by State",
       y = "Total Annual Loss of Bee Colonies")+
   theme( axis.text.x = element_text(angle = 60, hjust=0.8),
          axis.text=element_text(size=8),
          axis.title=element_text(size=10),
          legend.position="none",
          plot.title = element_text(lineheight=0.8, face="bold"))+
  scale_fill_brewer(palette = "Dark2")

```

##What really matters is the ratio of beekeepers to colonies, Right?

##But...
```{r}
bee %>% 
  mutate(coloniesPerKeeper = colonies/beekeepers) %>% 
  filter(year=="2016/17") %>% 
  ggplot(aes(x=coloniesPerKeeper, y= totAnnualLoss, color=state))+
  geom_jitter(shape=18, alpha=.8)+
  geom_smooth(method=lm, se=FALSE)+
  labs(title = " Ratio of Colonies to Beekeepers vs Total Annual Loss in 2016-2017",
       x = "Total Annual Loss",
       y = "Colonies  /  Beekeepers")+
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text.y = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=10),
          legend.position="none",
          plot.title = element_text(lineheight=.8, face="bold"))

```

##Furthermore...
```{r}
bee %>% 
  mutate(coloniesPerKeeper = colonies/beekeepers) %>% 
  filter(state == "Oklahoma" | state=="Iowa") %>% 
  ggplot(aes(x=totAnnualLoss, y= coloniesPerKeeper, color = year))+
  geom_jitter()+
  geom_smooth(method=lm, se=FALSE)+
  labs(title = "Total Anuual Loss vs Ratio of Colonies to Beekeepers",
       x = "Total Annual Loss",
       y = "Colonies : Beekeepers")+ 
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text.y = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=13),
          plot.title = element_text(lineheight=.8, face="bold"))+
  facet_grid(. ~ state)+
  scale_fill_brewer(palette = "OrRd")
```

-Iowa and Oklahoma are two states with some of the highest total annual loss.

##On the other hand...
```{r}
bee %>% 
  mutate(coloniesPerKeeper = colonies/beekeepers) %>% 
  filter(state == "California" | state== "Massachusetts") %>% 
  ggplot(aes(x=totAnnualLoss, y= coloniesPerKeeper, color = year))+
  geom_jitter()+
  geom_smooth(method=lm, se=FALSE)+
  labs(title = "Total Anuual Loss vs Ratio of Colonies to Beekeepers",
       x = "Total Annual Loss",
       y = "Colonies : Beekeepers")+ 
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text.y = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=13),
          plot.title = element_text(lineheight=.8, face="bold"))+
  facet_grid(. ~ state)+
  scale_fill_brewer(palette = "BuGn")
```

-California and Massachusetts typically have a low total annual loss. 

-We can see that once again, there is no linear regression line


##Conclusion

The number of beekeepers being hired is not related to rising rates colony collapse. But here are some other factors that are likely to be:

- Bullet 1
- Bullet 2
- Bullet 3

Source: <insert link to source here>.

