---
title: "Lab 2"
output: html_document
---
```{r}
sara <- 15
sara
```

```{r}
my_numeric <- 42
my_integer <- 2L #adding an L automatically denotes an integer
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i
```
```{r}
class(sara)
```
```{r}
## tp create new vector with 3 values
## vectors dont always have to be numeric
## c is like 'combine'
my_vector <- c(10, 20, 30)
my_vector
```
```{r}
## vectors can only hold data of the same type. cant combine to make a vectr with nmerics and characters
my_dates <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
my_dates
```
```{r}
# you can name the elements of the vector as well
#use the names() function included
names(my_dates) <- c("Good Day", "Bad Day", "Good Day", "Bad Day", "Good Day")
my_dates
# but this isn't really effiencient if you have to individually assign the elements one by one. 
```

```{r}
# a more effiecient way would be to use index
my_dates_GoodDays <- my_dates[c(1,3,5)]
my_dates_GoodDays
```

## Practice
Let's create a new numeric vector.
```{r}
new_vector <- 1:100
new_vector[new_vector == 15] #we can ask to select only the value 15
new_vector
```

We can evaluate all numerics in our vector to determine which has a value of 15. Try some other values to experiment.
```{r}
new_vector == 100
```

Now we are getting somewhere, because we can also use <, >, ==, <==, etc. Try another function to experiment.
```{r}
new_vector < 10
```

Using the `&` symbol allows us to refine our elements of interest.
```{r}
new_vector < 10 & new_vector > 2
```

If you use `[]` then you only get the values, not the logical evaluation of the entire vector.
```{r}
new_vector[new_vector < 10 & new_vector > 2]
new_vector
## the '&' operator means it must match the conditions. '|' is like 'or'
new_vector[new_vector <= 10 | new_vector >= 2]
```
```{r}
# commit is like a local save that doesnt overwrite other versions-never press ammend previous versions 
```

